\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-21 forwards-and-backwards.tex
.

Putting these together,
the level $u$ of an individual at $x$ in $\lp$ evolves according to:
$$
    \dot u
    =
    - \frac{\theta}{\lambda} u \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
$$
The factor of $\theta$ appears here because we have rescaled time.
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align} \label{eqn:b_limit}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } \theta \to \infty .
\end{align}

@@

Although we will \comment{later} show convergece of the lookdown process
defined by Definition~\ref{defn:lookdown_mgale}
to a stochastic process in which individuals retain their identities,
it is convenient to construct the lookdown process
in a way that explicitly represents the motion of each individual and their offspring.
This will then make obvious how linages move in the limit.

We will construct the process using a Ulam-Harris indexing scheme,
which associates the outcomes of each branching event with a label obtained
by appending to the previous label.
To begin, we assign each extant individual a label from $\IN_{\ge 0}$.
Suppose an individual with label $a$ and level $u$ reproduces,
and as a result there are two individuals, one with level $u$ and one with a new level $u_1 > u$.
The parent individual, previously labeled $a$, might be assigned either level.
After reproduction, we give a new label to \emph{only} the individual
that is given the new level $u_1$,
retaining the label $a$ for the individual with the old level $u$.
So, there is a unique label assigned at each birth event,
and it is assigned to the resulting individual with the higher level.
It would be possible to reconstruct individual identities along a line of descent
from this construction,
but we don't write this out.

Concretely, then: for each label $a$ in
$\labelspace = \bigcup_{k \ge 1} \IN_{\ge 1}^k$
let $\Pi_a$ be an independent Poisson process on $[0, \infty)^2 \times \IR_d \times \{0,1\}$.
The mean measure is a product of Lesbegue measure on the first three coordinates
and $q(0, \cdot)$ on the remainder.
For a point $(t, v, y, kappa)$ in $\Pi_a$,
first $t$ will determine possible birth times,
$v$ will determine levels of the offspring,
$y$ will give the spatial displacement of the offspring relative to the parent,
and $\kappa$ will be used to determine whether parent or offspring are assigned the new level.
Each label $a$ has a birth time $\tau_a$,
when it is first assigned,
and a death time $\sigma_a$, when its level first hits $\lambda$.
For any $\tau_a \le t \le \sigma_a$ we denote by $X_a(t)$ and $U_a(t)$ the spatial location and level
of the individual carrying label $a$ at time $t$.
Now, since we have defined labels so that the level does not jump,
$U_a$ satistifies, for $\tau_a \le t \le \sigma_a$,
\begin{equation*}
    \begin{split}
& U_a(t)
    =
    U_a(\tau_a) \\
&\qquad {}   
    + \int_{\tau_a}^{t}
    \left(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s)) \int_{\IR^d} r(y,\eta(s)) q(X_a(s),dy) U_a(s)^2
        -
        b(X_a(s),\eta(s)) U_a(s)
    \right)
    ds ,
\end{split}
\end{equation*}
and, of course, $\sigma_a = \inf\{t \ge \tau_a : U_a(t) > \lambda\}$.

Potential reproduction events occur at times $\tau_a + s$ for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ such successful event,
\comment{(need to sort out the "successful" qualifier)}
the new level is $aj$ (i.e., $a$ with the integer $j$ appended),
and the (potential) new level is
\begin{equation*}
    U_{aj}(\tau_a + s)
    =
    U_a(\tau_a + s)
    +
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(X_a(\tau_a + s-), \eta(\tau_a + s-))
        \int_{\IR^d} r(y, \eta(\tau_a + s-)) q(X_a(\tau_a + s-), dy)
    }.
\end{equation*}
If $U_{aj}(\tau_a+s) < \lambda$,
a new individual is produced with this level, label $aj$ and
location 
$$
    X_{aj}(\tau_a + s)
    =
    X_a(\tau_a+s-) + (1 - \kappa) y .
$$
If $\kappa=1$, then it is the original label that moves spatial locations, so
$$
    X_a(\tau_a+s)
    =
    X_a(\tau_a+s-) + \kappa y .
$$

>>>>

Putting these together,
the level $u$ of an individual at $x$ in $\lp$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \frac{\theta}{\lambda} u \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
\end{equation}
The factor of $\theta$ appears here because we have rescaled time.
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align} \label{eqn:b_limit}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } \theta \to \infty .
\end{align}

@@

Although we will \comment{(later)} show convergece of the lookdown process
defined by Definition~\ref{defn:lookdown_mgale}
to a stochastic process in which individuals retain their identities,
it is convenient to construct the lookdown process
in a way that explicitly represents the motion of each individual and their offspring.
This will then make obvious how linages move in the limit.

We will construct the process using a Ulam-Harris indexing scheme,
which associates the outcomes of each branching event with a label obtained
by appending to the previous label.
To begin, we assign each extant individual a label from $\IN_{\ge 0}$.
Suppose an individual with label $a$ and level $u$ reproduces,
and as a result there are two individuals, one with level $u$ and one with a new level $u_1 > u$.
The parent individual, previously labeled $a$, might be assigned either level.
We will track chains of descendant individuals forwards through time
following levels, rather than individuals, and will call this a \emph{line of descent}.
So, after reproduction, we give a new label to \emph{only} the individual
that is given the new level $u_1$,
retaining the label $a$ for the individual with the old level $u$.
So, there is a unique label assigned at each birth event,
and it is assigned to the resulting individual with the higher level.
It would be possible to reconstruct individual identities along a line of descent
from this construction,
but we don't write this out.

Concretely, then: for each label $a$ in
$\labelspace = \bigcup_{k \ge 1} \IN_{\ge 1}^k$
let $\Pi_a$ be an independent Poisson process on $[0, \infty)^2 \times \IR_d \times \{0,1\}$.
The mean measure is a product of Lesbegue measure on the first two coordinates,
$q(0, \cdot)$ on $\IR^d$, and $(\delta_0 + \delta_1)/2$ on $\{0, 1\}$.
It will also be convenient
to suppose that for each label $a$ we have an enumeration of the points in $\Pi_a$,
so we may refer to ``the $j^\text{th}$ point in $\Pi_a$'',
although the precise order of this enumeration is irrelevant.
If $(t, v, y, \kappa)$ is the $j^\text{th}$ point in $\Pi_a$,
then $t$ will determine possible birth times,
$v$ will determine levels of the offspring,
$y$ will give the spatial displacement of the offspring relative to the parent,
$\kappa$ will be used to determine whether parent or offspring are assigned the new level,
and the new label produced, if any will be $aj$,
i.e., the label $a$ with $j$ appended (so, if $a \in \IN_{\ge 0}^k$ then $aj \in \IN_{\ge 0}^{k+1}$).
Each label $a$ has a birth time $\tau_a$,
when it is first assigned,
and a death time $\sigma_a$, when its level first hits $\lambda$.
For any $\tau_a \le t \le \sigma_a$ we denote by $X_a(t)$ and $U_a(t)$ the spatial location and level
of the individual carrying label $a$ at time $t$.
Now, since we have defined labels so that the level does not jump,
$U_a$ satistifies~\eqref{eqn:dot_u} for $\tau_a \le t \le \sigma_a$, i.e.,
\begin{equation*}
    \begin{split}
& U_a(t)
    =
    U_a(\tau_a) \\
&\qquad {}   
    + \int_{\tau_a}^{t}
    \left(
        \frac{\theta}{\lambda} \gamma(X_a(s),\eta(s))
        \int_{\IR^d} r(z,\eta(s)) q(X_a(s),dz) U_a(s)^2
        -
        b(X_a(s),\eta(s)) U_a(s)
    \right)
    ds ,
\end{split}
\end{equation*}
and, of course, $\sigma_a = \inf\{t \ge \tau_a : U_a(t) > \lambda\}$.

Potential reproduction events occur at times $\tau_a + s$
for each point $(s, v, y, \kappa) \in \Pi_a$.
We say `potential' since if the resulting level is greater than $\lambda$,
the event does not happen.
If this is the $j^\text{th}$ point in $\Pi_a$,
the new label is $aj$ and the (potential) new level is
\begin{equation*}
    U_{aj}(\tau_a + s)
    =
    U_a(\tau_a + s)
    +
    \frac{
        v
    }{
        2 \lambda^{-1} \theta \gamma(X_a(\tau_a + s-), \eta(\tau_a + s-))
        r(X_a(\tau_a + s-) + y, \eta(\tau_a + s-))
    }.
\end{equation*}
If $U_{aj}(\tau_a+s) < \lambda$,
the new individual labeled $aj$ is produced,
and $\kappa$ determines which label is associated with the new location,
so
\begin{align*}
    X_{aj}(\tau_a + s)
    &=
    X_a(\tau_a+s-) + (1 - \kappa) y \\
    X_a(\tau_a+s)
    &=
    X_a(\tau_a+s-) + \kappa y .
\end{align*}
If $U_{aj}(\tau_a+s) \ge \lambda$, then $X_a$ is unchanged.
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.

Consequently, 
the \emph{line of descent}
leading from the initial generation
to any label $a \in \labelspace$
can be explicitly constructed as follows.
Supposed that $a = (a_1, a_2, \ldots, a_n) \in \IN^n$,
and for $1 \le k \le n$ let $(a)_k = (a_1, a_2, \ldots, a_k)$.
Then, let $\zeta_a(t)$ denote the counting process that records
which label is present in the line of descent at time $t$,
i.e., $\zeta_a(t) = \max\{k \ge 1 : \tau_{(a)_k} \le t\}$.
This then lets us define $X_a(t)$ all the way back to $t=0$
by setting $X_a(t) = X_{(a)_{\zeta_a(t)}}(t)$ (and, $U_a(t)$ similarly).
Now, we can explicitly write down the evolution of $U_a(t)$ and $X_a(t)$.

First, for each $1 \le k < n$
let $(s_k, v_k, y_k, \kappa_k)$ be the $a_{k+1}^\text{th}$ point in $\Pi_{(a)_k}$
(this is the sequence of jumps taken by the line of descent).
Note that $s_k = \tau_{(a)_k}$, and also set $s_0 = 0$.
Let $\Gamma_a$ by the point process ``experienced by the line of descent leading to $a$'',
which is for $A \subset [0,\infty)^2 \times \IR^d \times \{0,1\}$
$$
\Gamma_a(A)
=
    \sum_{k=1}^n
    \Pi_{(a)_k}\left(A \cap [s_{k-1}, s_k) \times [0, \infty) \times \IR^d \times \{0,1\}\right) .
$$
Then for $0 \le t \le \sigma_a$,
\begin{align*}
    X_a(t)
    &=
    X_{a_1}(0)
    +
    \sum_{k=1}^{\zeta_a(t) - 1}
    y_k
    +
    \int_{
        [0, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
    }
    \kappa y
    \Gamma_{a}(ds, dv, dy, d\kappa) .
\end{align*}

% Then for $0 \le t \le \sigma_a$,
% \begin{align*}
%     X_a(t)
%     &=
%     X_{a_1}(0)
%     +
%     \sum_{k=1}^{\zeta_a(t) - 1}
%     \left(
%     y_k
%     +
%     \int_{
%         [s_{k-1}, s_{k}) \times [0, \infty) \times \IR^d \times \{0, 1\}
%     }
%     \kappa y
%     \Pi_{(a)_k}(ds, dv, dy, d\kappa)
%     \right)
%     \\
%     &\qquad {}
%     +
%     \int_{
%         [s_{\zeta_a(t)}, t) \times [0, \infty) \times \IR^d \times \{0, 1\}
%     }
%     \kappa y
%     \Pi_{(a)_{\zeta_a(t)}}(ds, dv, dy, d\kappa) .
% \end{align*}
